Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2021-10-06 paper/SupMat.Rnw
finished revision for section 1-5

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}

@@

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}

@@

Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}

@@

\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\

@@

$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}

@@

Whereas $\olep_i^{[EC]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
more than just marginals, this quantity is still independent of the choice between our left- and right-tail-association 
cases because the bivariate random variables $(E_i^\nat , C_{i \bs i}^\nat )$ and 
$(E_i^\shp ,C_{i \bs i}^\shp)$ have had their tail association removed, and the only differences 
between $(E_i^{(l)},C_{i \bs i}^{(l)})$ and $(E_i^{(r)},C_{i \bs i}^{(r)})$ are in their tail association
properties (i.e., these two bivariate distributions have the same marginals and covariances).
In other words, $(E_i^\nat , C_{i \bs i}^\nat )$ is distributed in the same way for the left- and right-tail-association cases,
as is $(E_i^\shp ,C_{i \bs i}^\shp)$.

We now examine $\olep_i^{[E \nat C ]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
\begin{equation}
\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[E \nat C]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

>>>>

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))] - \E[\ln(1-\delta+\delta\exp(b_i^\shp-b_j^\shp))],\label{eq:decomp:contrib_EC_1}

@@

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]
, \label{eq:decomp:contrib_EC}
$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$ and $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent.
Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_EC})
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))],\label{eq:decomp:contrib_ATA_2}

@@

@@

\olep_j^{[E \nat C]} &= \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] - \E[r_j(E_j^\shp , C_{j \bs i}^\shp)]\\
&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
\label{eq:decomp:j:EChard}
&= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)]

@@

$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other.

@@

Whereas $\olep_i^{[E \nat C]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
more than just marginals, this quantity is still independent of the choice between our left- and right-tail-association 
cases because the bivariate random variables $(E_i^\nat , C_{i \bs i}^\nat )$ and 
$(E_i^\shp ,C_{i \bs i}^\shp)$ have had their tail association removed, and the only differences 
between $(E_i^{(l)},C_{i \bs i}^{(l)})$ and $(E_i^{(r)},C_{i \bs i}^{(r)})$ are in their tail association
properties (i.e., these two bivariate distributions have the same marginals and covariances).
In other words, $(E_i^\nat , C_{i \bs i}^\nat )$ is distributed in the same way for the left- and right-tail-association cases,
as is $(E_i^\shp ,C_{i \bs i}^\shp)$.

We now examine $\olep_i^{[EC ]}$, which, by (\ref{eq:decomp:contrib_ATA_2}), equals 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.
